The Gender Pay Gap in the General Social Survey

### 2023-08-31 Paper/paper.tex
proofread papper

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{???}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm]{geometry}

@@

%% citation setup
\usepackage{csquotes}

@@

\setlength{\droptitle}{-2em}
  \title{The Gender Pay Gap in the General Social Survey}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Statistical Learning}
  \author{Alex Amaguaya, Lea Bergmann}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{today}

@@

  \noindent\begin{minipage}[t]{0.3\textwidth}
  \IfLanguageName{english}{Deadline:}{Abgabefrist:}
  \end{minipage}
  \begin{minipage}[t]{0.7\textwidth}
  \hspace{1cm}tomorrow
  \end{minipage}

@@

  123456

@@

PhD Economics

@@

@@

  John Doe

  234567

  john.doe@web.de

  M.Sc. Economics

  \nth{4}

  Summer Term 2020

@@

The prediction of individuals' wages has been an issue in previous
literature. As wages are relevant to companies they want to know how
predict wages \autocite{Chakraborti}. Some papers make use of
classification techniques to predict wages \autocite{Chakraborti}. As we
have numeric values of real wages and not a categorical variable we do
not make use of classification techniques but prefer linear regression
techniques such as OLS and Lasso regressions as well as random forests.
We prefer these techniques as they handle categorical variables and
outliers well, which are features of our data \autocite{Cutler}. Wages
were already predicted using random forests by \autocite{Eichinger}.
They find that using random forests results in better predictions than
using linear models. Additionally, they show that gender barely effects
the wage predictions. Likewise, the boosting methods used by
\autocite{Chakrabarty2018} in salary level classification problems
showed better performance compared to a traditional algorithm such as
SVM. Furthermore, \autocite{BonaccoltoToepfer1612} suggests the use of a
penalization approach for high dimensional problems in his research
about Gender wage inequality.

@@

\begin{itemize}
\item
  Linear Regression
\item
  Logistic Regression
\item
  Lasso, Ridge and Elastic Net Regression
\item
  Gradient Boosting
\item
  Random Forest
\end{itemize}

@@

Before modeling we conduct a data preparation process. The initial data
set has 54,842 observations and the outcome variable \texttt{realrinc}
has 38.55\% of missing values. The imputation of the target variable is
discarded to avoid generating a bias in the modeling and therefore we
decided to eliminate these records from the original data set. After
this, the dataset contains 33,702 observations. Furthermore, it is
impossible to deduce the occupation for some observations
(\texttt{occ10\ =\ 9997}) and we remove these from the data set. At the
end, the final data set has 33,244 observations, and we use it for the
modeling process.

This data set still had some missing values for the regressors, e.g.,
number of children, age, marital status, degree level, and others. We
use the \texttt{mice} package with five variables for the imputation.
The features \emph{age, occupational prestige and number of children}
are imputed using the pmm (\texttt{predictive\ mean\ matching}) method.
We use the polytomous logistic regression method for the factor
variables (\emph{education and marital status}). The imputation process
generates 6 different data sets. We apply the estimation of the mean and
the majority vote to aggregate the data sets into a single data set. The
mean estimation is applied for the numeric variables and majority vote
for the categorical variables..

@@

After the imputation process, we create the interaction variables.
First, we estimate some interaction features using only the numerical
variables. Then, we transform the numerical variables into categorical
variables by dividing them into different ranges (e.g., age between 18
and 30, age between 31 and 50, etc.). We transform these new categorical
variables and the initial categorical variables (\emph{marital status,
education, etc.}) into dummy variables. And finally, the interaction
variables are estimated with the group of dummy variables mentioned
above. The inclusion of the interaction of the numeric variables are
validate using a linear model, and the results show that the p- values
of the parameters for the interactions are less than 0.05.

@@

Furthermore, we estimate a correlation matrix using Spearman's rank
correlation coefficient with some numeric variables in order to support
the previous results. The correlation results are shown in the following
figure.

@@

After the data preparation and feature engineering processes, we
construct a final data set with 33,244 observations and 421 features.
Due to the large number of interaction variables, we decided to reduce
them and use only the interaction between a group of variables
(\emph{occrecode, educcat and prestg10}). After reduction, we develop
the modeling process with 100 variables (numerical and interaction
variables). The following figure shows the overview of the group of
variables that were estimated and used in the models.

@@

Hence, the coefficient estimates in a ridge regression minimize the sum
of the residual sum of squares (RSS) and the sum of squared coefficients
multiplied with a tuning parameter \(\lambda\). We estimate \(\lambda\)
using cross-validation.

@@

According to the literature review, tree-based models perform
considerably well in range in predicting income ranges (classification
problems). For this reason, we use the random forest algorithm for
modeling. This experiment use the income without logarithm
transformation, the initial features and some interaction variables.

@@

\NormalTok{instance.rf }\OtherTok{=} \FunctionTok{tune}\NormalTok{(}
  \AttributeTok{tuner =}\NormalTok{ tnr\_rdgrid\_search,}
  \AttributeTok{task =}\NormalTok{ task,}
  \AttributeTok{learner =}\NormalTok{ learner,}
  \AttributeTok{resampling =}\NormalTok{ rsmp\_cv3,}
  \AttributeTok{measures =}\NormalTok{ msr\_ce,}
  \AttributeTok{term\_evals =} \DecValTok{150}\NormalTok{,}
  \AttributeTok{store\_models =} \ConstantTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Due to the number of hyper parameters in random forest, we use the
\texttt{lightgbm} and \texttt{mlr3} packages to model and tune the hyper
parameter in a time-optimal way. After many attempts, we found the best
hyper parameters that produced the lowest error. The hyper parameters
tuned were \texttt{max\_depth}, \texttt{num\_leaves},
\texttt{num\_iterations} and others. In addition, we used a random
search approach in the tuning process with a cross-validation of 3. Once
we found the model along with the hyper parameters with lowest error in
the cross-validation, we proceeded to evaluate this model on the test
data set and obtained a RMSE of 31,745.54. The hyper parameters of the
best model are shown in the following table.

{[}best hyper parameter of RF{]}

In addition, we can analyze the importance of the features within the
modeling process using random forest, and these results are shown in the
following table.

{[}feature importance of RF{]}

@@

Gradient Boosting also has many hyper parameters like random forest and
we decided to apply a similar approach to the previous algorithm to find
the best model with the hyper parameters (random search approach in the
tuning process with a cross-validation of 3). Furthermore, the
experiments with Gradient Boosting use the target variable without
logarithm transformation.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tnr\_rdgrid\_search }\OtherTok{=} \FunctionTok{tnr}\NormalTok{(}\StringTok{"random\_search"}\NormalTok{, }\AttributeTok{batch\_size =} \DecValTok{10}\NormalTok{)}
\NormalTok{rsmp\_cv3 }\OtherTok{=} \FunctionTok{rsmp}\NormalTok{(}\StringTok{"cv"}\NormalTok{, }\AttributeTok{folds =} \DecValTok{3}\NormalTok{)}
\NormalTok{msr\_ce }\OtherTok{=} \FunctionTok{msr}\NormalTok{(}\StringTok{"regr.rmse"}\NormalTok{)}
\NormalTok{learner.gb }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
                 \AttributeTok{boosting =} \StringTok{"gbdt"}\NormalTok{,}
                 \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
                 \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{6}\NormalTok{,}\DecValTok{8}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{8}\NormalTok{,}\DecValTok{11}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
                 \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}\FloatTok{0.2}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
                 \AttributeTok{learning\_rate =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.03}\NormalTok{, }\FloatTok{0.04}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
                 \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)), }\CommentTok{\#40}
                 \AttributeTok{lambda\_l1 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.4}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.1}\NormalTok{)),}
                 \AttributeTok{lambda\_l2 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.30}\NormalTok{, }\FloatTok{0.33}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
                 \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
                 \AttributeTok{num\_threads =} \DecValTok{5}
\NormalTok{)}

\NormalTok{instance.gb }\OtherTok{=} \FunctionTok{tune}\NormalTok{(}
  \AttributeTok{tuner =}\NormalTok{ tnr\_rdgrid\_search,}
  \AttributeTok{task =}\NormalTok{ task,}
  \AttributeTok{learner =}\NormalTok{ learner.gb,}
  \AttributeTok{resampling =}\NormalTok{ rsmp\_cv3,}
  \AttributeTok{measures =}\NormalTok{ msr\_ce,}
  \AttributeTok{term\_evals =} \DecValTok{170}\NormalTok{,}
  \AttributeTok{store\_models =} \ConstantTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We tried many experiments and found the best hyper parameters that
produced the lowest error. In Gradient Boosting modeling we adjusted the
hyper parameters such as \texttt{max\_depth}, \texttt{num\_leaves},
\texttt{min\_data\_in\_leaf}, \texttt{feature\_fraction} and others.
Once the best model was found, it obtained an RMSE of 29,207.55 on the
test dataset. The hyper parameters of the best model are shown in the
following table.

{[}best hyper parameter of GB{]}

Furthermore, the following table shows the importance of the features
within the modeling process using Gradient Boosting.

{[}feature importance of GB{]}

These results show that in tree-based methods, the Gradient Boosting
performance overcomes Random Forest, in other words, the error of
Gradient Boosting is approximately 8\% lower than the error of Random
Forest. In addition, the errors of these models are larger than the
linear models.

>>>>

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{???}
\fi
% use upquote if available, for straight quotes in verbatim environments
%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm]{geometry}

@@

%% citation setup
%\usepackage{csquotes}

@@

\setlength{\droptitle}{-2em}
  \title{The Gender Pay Gap in the General Social Survey}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Statistical Learning}
  \author{Alex Amaguaya, Lea Bergmann}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{31.08.2023}

@@

  \noindent\begin{minipage}[t]{0.3\textwidth}
  \IfLanguageName{english}{Deadline:}{Abgabefrist:}
  \end{minipage}
  \begin{minipage}[t]{0.7\textwidth}
  \hspace{1cm}31.08.2023
  \end{minipage}

@@

  3104233

@@

Ph.D. Economics

@@

  Summer Term 2028

  Alex Amaguaya

  242833

  alex.amaguaya@tu-dortmund.de

  M.Sc. Econometrics

  \nth{2}

@@

@@

The prediction of individuals' wages has been an issue in previous
literature. As wages are relevant to companies they want to know how to
predict wages \autocite{Chakraborti}. Some papers make use of
classification techniques to predict wages \autocite{Chakraborti}. As we
have numeric values of real wages and not a categorical variable we do
not make use of classification techniques but prefer linear regression
techniques such as OLS and Lasso regressions as well as random forests.
We prefer these techniques as they handle categorical variables and
outliers well, which are features of our data \autocite{Cutler}. Wages
were already predicted using random forests by \autocite{Eichinger}.
They find that using random forests results in better predictions than
using linear models. Additionally, they show that gender barely effects
the wage predictions. Likewise, the boosting methods used by
\autocite{Chakrabarty2018} in salary level classification problems
showed better performance compared to a traditional algorithm such as
support vector machines (SVM). Furthermore,
\textcite{BonaccoltoToepfer1612} suggest the use of a penalization
approach for high dimensional problems in his research about gender wage
inequality.

@@

\begin{itemize}
\item
  Linear Regression
\item
  Logistic Regression
\item
  Lasso, Ridge and Elastic Net Regression
\item
  Random Forest
\item
  Gradient Boosting
\item
  Generalized Additive Model
\end{itemize}

@@

Before modeling we conduct a data preparation process. The initial data
set has 54,842 observations and the outcome variable \texttt{realrinc}
has 38.55\% of missing values. The imputation of the target variable is
discarded to avoid generating a bias in the modeling and therefore we
eliminate these records from the original data set. After this, the data
set contains 33,702 observations. Furthermore, it is impossible to
deduce the occupation for some observations (\texttt{occ10\ =\ 9997})
and we remove these from the data set. Finally, the data set has 33,244
observations and we use it for the modeling process.

This data set still has some missing values for the regressors, e.g.,
number of children, age, marital status, degree level, and others. Thus,
we use the \texttt{mice} package with five variables for the imputation
process. The features \emph{age, occupational prestige and number of
children} are imputed using the predictive mean matching (pmm) method,
and we use the polytomous logistic regression method for the factor
variables (\emph{education and marital status}). As we use a value of 6
in the \texttt{m} parameter of the \texttt{mice} function, the
imputation process generates 6 different data sets. This parameter
refers to the number of imputed data sets. After this, we apply the
calculation of the mean and the majority vote to aggregate the data sets
into one. Mean estimation is applied for numerical variables and
majority voting for categorical variables

@@

After the imputation process, we create interaction variables. First, we
estimate some interaction features using only the numerical variables.
Then, we transform the numerical variables
(\texttt{age,\ childs\ and\ prestg10}) into categorical variables by
separating them into ranges. The ranges created of these variables were:
\texttt{age\ between\ 18\ and\ 30,\ age\ between\ 31\ and\ 50},
\texttt{age\ greater\ than\ 50;\ childs\ between\ 0\ and\ 2,\ childs\ between\ 3\ and\ 5,\ childs\ more\ than\ 5;\ prestg10\ between\ 16\ and\ 30,\ prestg10\ between\ 31\ and\ 50,\ prestg10\ greater\ than\ 50}.
After this, we transform these new categorical variables and the initial
categorical variables (\texttt{marital\ status}, \texttt{education},
etc.) into dummy variables. Finally, we estimate the interaction
variables with the group of dummy variables mentioned above (the
interaction consists of the multiplication between two dummy variables).
The inclusion of the interactions of the numeric variables is validate
using a linear model, and the results show that the p-values of the
parameters for the interactions are below 0.05.

@@

Furthermore, we estimate a correlation matrix using Spearman's rank
correlation coefficient with some numeric variables in order to support
the previous results. The correlation results are exposed in figure 2,
and show that \texttt{prestg10}, \texttt{age} and
\texttt{interaction\ between\ age\ and\ prest10} have the highest
positive correlation with \texttt{realrinc}, meanwhile the dummy
variable \texttt{female} has a moderate negative correlation with the
target variable.

@@

As the correlation assesses the linear relationship between two
variables and it only uses numerical features, we additionally estimate
the Predictive Power Score (PPS). It overcomes these limitations and
allows us to assess the predictive power of a regressor on a target
variable. This measure is a normalized metric and its values are between
0 (no predictive power) and 1 (perfect predictive power). The PPS is
calculated by comparing the performance of a model that uses a regressor
\texttt{x} to predict target variable \texttt{y} versus the performance
of a base model (naive). When the target variable is a numerical feature
it is common to use a Decision Tree Regressor to predict \texttt{y}
based on \texttt{x}, while in categorical cases a Decision Tree
Classifier is used. On the other hand, base models are often defined as
mean prediction or random or modal prediction for regression and
classification problems, respectively. The PPS results are presented in
figure 3, and show that \texttt{prestg10}, \texttt{wrkstat},
\texttt{educcat} and \texttt{occrecode} have the highest predictive
power of \texttt{realrinc} relative to the other features. These results
provide some insights in the data exploration process, but we must
beware of the limitations of this approach.

\begin{figure}
\centering
\includegraphics{includes/pps.png}
\caption{PPS Results}
\end{figure}

After the data preparation and feature engineering processes, we have a
final data set with 33,244 observations and 421 features. Due to the
large number of interaction variables, we reduce them and use only the
interaction between a group of variables (\texttt{occrecode},
\texttt{educcat} and categorical variable of \texttt{prestg10}). After
this, we develop the modeling process with 100 variables (numerical and
interaction variables). Figure 4 shows the overview of the group of
estimated variables that are used in the models.

@@

Hence, the coefficient estimates in a ridge regression minimize the sum
of the residual sum of squares (RSS) and the sum of squared coefficients
multiplied with a tuning parameter \(\lambda\). We estimate \(\lambda\)
using CV.

@@

According to the literature review, tree-based models perform
considerably well in predicting income ranges (classification problems)
\autocite{Chakrabarty2018}. For this reason, we use the random forest
algorithm for modeling. For this method we use income without logarithm
transformation, initial variables and some interaction variables. Due to
the amount of hyperparameters in random forest, we use the
\texttt{lightgbm} and \texttt{mlr3} packages to model and tune the
hyperparameter in a time-optimal way. We tune the hyperparameters
\texttt{max\_depth}, \texttt{num\_leaves}, \texttt{num\_iterations},
etc. to find the best hyperparameters that produce the lowest error. The
search space of the hyperparameters is showed below. In addition, we use
a random search approach in the tuning process with a CV of 3 folds.

@@

\NormalTok{instance.rf }\OtherTok{=} \FunctionTok{tune}\NormalTok{(}
  \AttributeTok{tuner =}\NormalTok{ tnr\_rdgrid\_search,}
  \AttributeTok{task =}\NormalTok{ task,}
  \AttributeTok{learner =}\NormalTok{ learner,}
  \AttributeTok{resampling =}\NormalTok{ rsmp\_cv3,}
  \AttributeTok{measures =}\NormalTok{ msr\_ce,}
  \AttributeTok{term\_evals =} \DecValTok{300}\NormalTok{,}
  \AttributeTok{store\_models =} \ConstantTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

After identifying the model along with the hyperparameters with lowest
error in CV, we proceed to evaluate this model on the test data set and
obtain an RMSE of 31,745.54. The hyperparameters of the best model are
shown in figure 5.

\begin{figure}
\includegraphics[width=0.5\linewidth]{includes/RF_hyper} \caption{Best Hyperparameters for Random Forest}\label{fig:unnamed-chunk-1}
\end{figure}

In addition, we analyze the features importance within the modeling
process using random forest. These results are exposed in the figure 6.

\begin{figure}
\centering
\includegraphics{includes/RF_importance.png}
\caption{Top 20 of Features Importance for Random Forest}
\end{figure}

Unlike the linear models presented previously, random forest does not
predict negative values when logarithmic transformation is not used.
Still, we additionally use a random forest with the logarithmic
transformation in the target variable and keep the same configuration
(CV of 3 folds and random search approach in the tuning process).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{learner.wlog }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
              \AttributeTok{boosting =} \StringTok{"rf"}\NormalTok{,}
              \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
              \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{7}\NormalTok{,}\DecValTok{12}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
              \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{120}\NormalTok{,}\DecValTok{140}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
              \AttributeTok{max\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{45}\NormalTok{,}\DecValTok{60}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
              \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{50}\NormalTok{,}\DecValTok{90}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
              \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{20}\NormalTok{,}\DecValTok{35}\NormalTok{,}\DecValTok{1}\NormalTok{)), }
              \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{35}\NormalTok{,}\DecValTok{50}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
              \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.4}\NormalTok{,}\FloatTok{0.7}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
              \AttributeTok{bagging\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.6}\NormalTok{,}\FloatTok{0.9}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
              \AttributeTok{bagging\_freq =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{3}\NormalTok{,}\DecValTok{12}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
              \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.8}\NormalTok{,}\FloatTok{0.9}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
              \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
              \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
              \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
              \AttributeTok{num\_threads =} \DecValTok{5}\NormalTok{,}
              \AttributeTok{seed =} \DecValTok{123}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The best model using the logarithmic transformation has different
hyperparameters relative to the previous experiment and obtains an RMSE
of 34,293.33 on the test data set. Thus, the error of this model is
greater than the best model without the transformation (31,745.54).

@@

Gradient Boosting has a similar amount of hyperparameters as random
forest and we use a similar configuration as the previous algorithm to
find the best model with the hyperparameters (random search approach in
the tuning process with 3-fold CV). The first model with Gradient
Boosting uses the target variable without logarithmic transformation.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tnr\_rdgrid\_search }\OtherTok{=} \FunctionTok{tnr}\NormalTok{(}\StringTok{"random\_search"}\NormalTok{, }\AttributeTok{batch\_size =} \DecValTok{10}\NormalTok{)}
\NormalTok{rsmp\_cv3 }\OtherTok{=} \FunctionTok{rsmp}\NormalTok{(}\StringTok{"cv"}\NormalTok{, }\AttributeTok{folds =} \DecValTok{3}\NormalTok{)}
\NormalTok{msr\_ce }\OtherTok{=} \FunctionTok{msr}\NormalTok{(}\StringTok{"regr.rmse"}\NormalTok{)}
\NormalTok{learner.gb }\OtherTok{=} \FunctionTok{lrn}\NormalTok{(}\StringTok{"regr.lightgbm"}\NormalTok{,}
                 \AttributeTok{boosting =} \StringTok{"gbdt"}\NormalTok{,}
                 \AttributeTok{objective =} \StringTok{"regression"}\NormalTok{,}
                 \AttributeTok{max\_depth =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{num\_leaves =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{6}\NormalTok{,}\DecValTok{8}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{min\_data\_in\_leaf =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{min\_data\_in\_bin =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{8}\NormalTok{,}\DecValTok{11}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{feature\_fraction =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.3}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
                 \AttributeTok{feature\_fraction\_bynode =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}\FloatTok{0.2}\NormalTok{,}\FloatTok{0.1}\NormalTok{)),}
                 \AttributeTok{learning\_rate =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.03}\NormalTok{, }\FloatTok{0.04}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
                 \AttributeTok{num\_iterations  =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{39}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{1}\NormalTok{)),}
                 \AttributeTok{lambda\_l1 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.4}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.1}\NormalTok{)),}
                 \AttributeTok{lambda\_l2 =} \FunctionTok{to\_tune}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\FloatTok{0.30}\NormalTok{, }\FloatTok{0.33}\NormalTok{, }\FloatTok{0.01}\NormalTok{)),}
                 \AttributeTok{convert\_categorical =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{force\_col\_wise =} \ConstantTok{TRUE}\NormalTok{,}
                 \AttributeTok{verbose =} \DecValTok{1}\NormalTok{,}
                 \AttributeTok{num\_threads =} \DecValTok{5}
\NormalTok{)}

\NormalTok{instance.gb }\OtherTok{=} \FunctionTok{tune}\NormalTok{(}
  \AttributeTok{tuner =}\NormalTok{ tnr\_rdgrid\_search,}
  \AttributeTok{task =}\NormalTok{ task,}
  \AttributeTok{learner =}\NormalTok{ learner.gb,}
  \AttributeTok{resampling =}\NormalTok{ rsmp\_cv3,}
  \AttributeTok{measures =}\NormalTok{ msr\_ce,}
  \AttributeTok{term\_evals =} \DecValTok{300}\NormalTok{,}
  \AttributeTok{store\_models =} \ConstantTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Our best model uses a combination of hyperparamters that produces the
least error. We adjust some hyperparameters such as \texttt{max\_depth},
\texttt{num\_leaves}, \texttt{min\_data\_in\_leaf},
\texttt{feature\_fraction} and others. We evaluate the best model with
the test data set and obtain an RMSE of 29,207.55. The hyperparameters
of the best model are shown in figure 7.

\begin{figure}
\includegraphics[width=0.5\linewidth]{includes/GB_hyper} \caption{Best Hyperparameters for Gradient Boosting}\label{fig:unnamed-chunk-2}
\end{figure}

Gradient Boosting also provides information on the importance of
features during the modeling process in this experiment. Figure 8 shows
the 20 most important variables according to the Gini metric.
\texttt{Gender} is after work status the most important variable, giving
additional evidence for a gender pay gap.

\begin{figure}
\centering
\includegraphics{includes/GB_importance.png}
\caption{Top 20 of Features Importance for Gradient Boosting}
\end{figure}

Furthermore, we estimate a gradient boosting model using the logarithm
transformation on the target variable.
